|

### 2025-12-03 manuscript/pre_print.tex
Add Nguo et al (2025) reference

Therefore, to estimate and describe the magnitude of difference in REE
between women with and without PCOS, we completed a systematic review
and meta-analysis of studies reporting REE in these populations.

@@

This systematic review and meta-analysis was pre-registered on PROSPERO
(\href{https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434}{CRD42024601434})
initially on the \(3^{rd}\) of December 2024 and performed in accordance
with the Preferred Reporting Items for Systematic Reviews and
Meta-Analyses (PRISMA) statement guidelines\textsuperscript{30}. The
PRISMA flow diagram reported below (Figure~\ref{fig-prisma}) was
produced using the \texttt{PRISMA2020} R package and Shiny
app\textsuperscript{31}. The primary aim of this review was to examine
the descriptive question ``Does resting energy expenditure (REE) differ
between women with and without polycystic ovary syndrome (PCOS)?''. We
summarise and describe the studies in addition to quantitatively
synthesising their results via meta-analysis.

@@

The condition being studied was PCOS and we included observational
cross-sectional design studies, in addition to intervention studies
where REE was reported for the population (and if present, the
comparator i.e., women without PCOS) condition of interest. For clarity,
studies of any design were included if they reported the REE using the
methods indicated for a sample of adult women with PCOS and who were
otherwise healthy. This included both studies with and without samples
of healthy control women without PCOS. As detailed in the statistical
analysis section below, a Bayesian model with informative priors based
on normative data for REE in healthy women without PCOS was included to
provide control information indirectly where this was missing. The use
of such priors is an efficient tool for incorporating historical
information on a particular population in a conservative
manner\textsuperscript{32}.

@@

For each arm, and observation time point if multiple observations
reported (e.g., before and after an intervention), depending on what was
reported by the authors we extracted the means, medians, standard
deviations, standard errors, lower and upper range values, and
interquartile range for the unadjusted and/or body mass adjusted and/or
fat free mass adjusted REE values. Where REE values adjusted for body
mass and/or fat free mass were reported we used the reported body mass
and/or fat free mass mean values for that arm to convert them to
unadjusted REE values (i.e., multiplied them by body mass and/or fat
free mass mean values). Where means and/or standard deviations were
missing the latter were either calculated from standard errors and
sample size, or all both were estimated from lower and upper range,
interquartile range, median, and sample size depending on the available
information using the methods of Wan et al.\textsuperscript{33}.
Further, where missing, height/body mass/BMI where estimated based on
the reported means. The units of measurement for which REE was extracted
and all REE values were converted to kcal/day. In one
case\textsuperscript{34} REE was reported relative to body mass and the
unadjusted values were no longer available (confirmed by the authors).
As such, in this case we used the mean body mass to convert back to
estimated REE unadjusted.

@@

During data extraction it was noted that several studies from the same
lab/research group\textsuperscript{13,35--37} contained a number of
discrepancies that seemed to be possible reporting errors. This
included, based on taking the authors results as written, standard
errors that implied impossible or at least incredibly unlikely standard
deviations, and discrepancies in sample size reporting throughout for
most variables without explanation or where this was explained the
sample sizes were discrepant with the text Further, data was not
reported for the healthy control women without PCOS in three of the
studies\textsuperscript{35--37}, and REE was reported as an ``adjusted''
value whereby
\(\textrm{REE}_{adjusted}=\textrm{REE}_{group~mean}+(\textrm{REE}_{adjusted}−\textrm{REE}_{predicted})\)
and the \(\textrm{REE}_{predicted}\) was obtained by substituting the
individual lean body mass, fat mass, gender, and age in the linear
regression equation generated by the data of all patients. In
correspondence with the senior author we were unable to clarify the
reporting discrepancies as the person responsible for the data/results
was no longer contactable. The original data were also no longer
available and so we could not calculated the unadjusted REE.

@@

Statistical analysis of the data extracted was be performed in R, (v
4.3.3; R Core Team, https://www.r-project.org/) and RStudio (v
2023.06.1; Posit, https://posit.co/). All code utilised for data
preparation, transformations, analyses, plotting, and reporting are
available in the corresponding GitHub repository
\url{https://github.com/jamessteeleii/pcos_ree_meta}. We cite all
software and packages used in the analysis pipeline using the
\texttt{grateful} package\textsuperscript{38} which can be seen here:
\url{https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf}.
The statistical analysis plan was linked in our pre-registration
(PROSPERO:
\href{https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434}{CRD42024601434})
and available at the accompanying GitHub repository. Any deviations from
the pre-registration are noted below.

Given our research question our analysis was aimed at parameter
estimation\textsuperscript{39} within a Bayesian meta-analytic
framework\textsuperscript{40}. For all analyses model parameter
estimates and their precision, along with conclusions based upon them,
are interpreted continuously and probabilistically, considering data
quality, plausibility of effect, and previous literature, all within the
context of each model. The \texttt{renv} package\textsuperscript{41} was
used for package version reproducibility and a function based analysis
pipeline using the \texttt{targets} package\textsuperscript{42} was
employed (the analysis pipeline can be viewed by downloading the R
Project and running the function \texttt{targets::tar\_visnetwork()}).
Effect sizes and their variances were all calculated using the
\texttt{metafor} packages \texttt{escalc()}
function\textsuperscript{43}. The main package
\texttt{brms}\textsuperscript{44} was used in fitting all the Bayesian
meta-analysis models. Prior and posterior draws were taken using
\texttt{marginaleffects}\textsuperscript{45} and
\texttt{tidybayes}\textsuperscript{46} packages. All visualisations are
created using \texttt{ggplot2}\textsuperscript{47}, \texttt{tidybayes},
and the \texttt{patchwork}\textsuperscript{48} packages.

@@

We adopted an arm-based multiple condition comparison (i.e., network)
type model given that the studies included had arms of women with PCOS
both with, and without, a non-PCOS control arm\textsuperscript{49}, and
also in some cases multiple observations of REE in the different arms
included in the study (for example, where an intervention was conducted
and pre- and post-intervention REE was reported). In typical
contrast-based meta-analyses data is limited to the effect sizes for
paired contrasts between arms and thus studies that include both arms
(i.e., relative effects between non-PCOS control vs PCOS arms); however,
in arm-based analyses the data are the absolute effects within each arm
and information is borrowed across studies to enable both within
condition absolute, and between condition relative contrasts to be
estimated. We made use of historical information regarding REE in
healthy control women without PCOS by setting informative priors based
on meta-analysis of large scale studies reporting normative data for REE
in this population. This was included to provide indirect control
information where it was missing from particular studies. The use of
historical priors like this is an efficient tool to incorporate
historical information about a particular population in a conservative
manner in meta-analyses\textsuperscript{32}. From this model we focus on
reporting the global grand mean estimate for the fixed between condition
relative contrast for non-PCOS control vs PCOS arms as our primary
estimand of interest (i.e., \(\beta_1\) in both mean and standard
deviation models). We examined both raw mean REE (i.e., the absolute
mean REE in kcals per day for each arm) in addition to the between
person standard deviation in REE (i.e., the absolute standard deviation
in REE in kcals per day for each arm). Both models were multilevel in
that they included nested random intercepts for both study and arm
within study. In addition, and in deviation from our pre-registration,
we also included lab as a random intercept as in some cases we had
multiple studies from the same lab or research group. Lastly, the
inclusion of a random intercept for each effect size was accidentally
omitted from our pre-registration, and so this is also included in the
model.

@@

\[
\small
\begin{aligned}
\beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]
\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]
\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{aligned}
\] where the prior for \(\beta_{0}\), which corresponded to the model
intercept and mean REE in the control condition, was set based on
meta-analysis of the mean REEs for women from two large studies of
healthy people\textsuperscript{50,51} though set with a conservative
degrees of freedom for the \(\textrm{Student-}t\) distribution.The
random intercept \(\sigma^2_{0,study}\) was set similarly to this. The
prior for the fixed effect \(\beta_{1}\), reflecting the difference
between control and PCOS conditions was set based on a wide range of
possible values considering the minimum and and maximum values of the
ranges reported in the two studies noted (i.e., 2492 - 908 = 1584). We
then set a prior that permits values approximately across this range of
values with the majority of it's mass centred around zero. The remaining
random effects were set based on the default weakly regularising priors
for \texttt{brms} and scaled to the expected response values using a
\(\textrm{half-student-}t\) distribution with 3 degrees of freedom and
\(\mu=0\), and both correlation matrices \(\textrm{corr}_{lab}\) and
\(\textrm{corr}_{study}\) were set with an \(\textrm{LKJcorr(1)}\)
distribution.

@@

where the prior for \(\beta_{0}\), which corresponded to the model
intercept and standard deviation of REE in the control condition, was
again set based on meta-analysis of the standard deviation of REEs for
women from two large studies of healthy people\textsuperscript{50,51}
though set with a conservative degrees of freedom for the
\(\textrm{Student-}t\) distribution.The random intercept
\(\sigma^2_{0,study}\) was set similarly to this. The prior for the
fixed effect \(\beta_{1}\), reflecting the difference between control
and PCOS conditions was set based on a wide range of possible values.
Given that in many cases of variables in the field there is an
approximate relationship of \textasciitilde1 for the natural logarithm
of the standard deviation conditioned upon the natural logarithm of the
mean\textsuperscript{52} we set this prior to reflect the range of
differences on the on the log scale (i.e., \(\log(1584)\)). We then set
a prior that permits values approximately across this range of values
with the majority of it's mass centred around zero. The prior for the
fixed effect \(\beta_{2}\), reflecting the relationship between the
natural logarithm of the mean REE with the natural logarithm of the
standard deviation of REE, was set it to be weakly informative centred
on zero with a wide scale to indicate uncertainty in this outcome
specifically despite the typical relationship close to \textasciitilde1.
Priors for the measurement error of the natural logarithm of the mean
REE estimate \(\tilde{m}\) were again based upon meta-analysis of the
aforementioned studies, though measurement error has to be positive, as
does the corresponding standard deviation of this error, so we set these
to conservative wide \(\textrm{half-student-}t\) distributions. The
remaining random effects were set based on the default weakly
regularising priors for \texttt{brms} and scaled to the expected
response values using a \(\textrm{half-student-}t\) distribution with 3
degrees of freedom and \(\mu=0\), and both correlation matrices
\(\textrm{corr}_{lab}\) and \(\textrm{corr}_{study}\) were set with an
\(\textrm{LKJcorr(1)}\) distribution.

@@

As noted in the section above, \emph{``Studies with possible reporting
errors''}, we opted to conduct analysis with and without the inclusion
of four studies with possible reporting errors we could not
resolve\textsuperscript{13,35--37}. Thus the main models described above
were run with and without these studies, the main results are presented
without them and the sensitivity results with them are reported
separately below.

As a further sensitivity analysis, given the inclusion of some studies
with interventions in women with PCOS reporting REE at multiple
timepoints such as mid- and post-intervention\textsuperscript{34,53,54}
but lacking of control women without PCOS, we opted to also conduct
sensitivity analysis excluding these and only examining their baseline
results in addition to the cross-sectional studies. As such, the main
models described above were run without the follow-up (i.e., mid- or
post-intervention timepoints) from these studies and only using the
baseline results in addition to other cross-sectional studies.

@@

Note, the numbers in this section exclude the four studies previously
noted with possible reporting issues that were unresolved. These studies
are however summarised in fully in the descriptive characteristics table
in the online supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table} where some of the possible reporting errors are seen in the
standard deviations reported or calculated e.g., the standard deviation
calculated for body mass of women with PCOS in Saltamavros et
al.\textsuperscript{36}. The PRISMA flow diagram
(Figure~\ref{fig-prisma}) also includes all studies identified including
the four with possible errors.

@@

Studies were carried out in multiple countries: Brazil (k = 3) ,and USA
(k = 3 studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK
(all k = 1 study). The four studies with noted reporting errors were
carried out in Greece. A total of 11 studies used the Rotterdam criteria
(or a modified version there of) for diagnosing
PCOS\textsuperscript{16,17,34,53--60}. One study used the 1990 National
Institutes of Health criteria\textsuperscript{61}, and two
studies\textsuperscript{18,55} diagnosed PCOS via the presence of
oligomenorrhea or amenorrhea alongside additional criteria including
plasma androgen levels, hirsutism or polycystic ovaries on ultrasound
scanning. All but one study\textsuperscript{61}, which used doubly
labelled water, measured REE using indirect calorimetry. The specific
devices reported by these studies are included in the online
supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}). We originally considered in our pre-registration that, given
sufficient data, we would compare sub groups of women with PCOS who did
and did not have accompanying insulin resistance. However, based on the
metabolic health variables reported in studies with this information
(see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}) and, where available, considering primary criteria of either
homeostatic model assessment of insulin resistance (HOMA-IR) \(\geq2.5\)
or secondary criteria including fasting insulin \(>12\mu \textrm{U/mL}\)
or fasting glucose \(\geq100\textrm{mg/dL}\), all groups of women with
PCOS in the included studies would be considered to have insulin
resistance. Mean age of women with PCOS in these studies ranged from 23
to 33 which was similar to the control women without PCOS ranging from
23 to 30. Across those studies where BMI was reported or it was possible
to estimate the mean BMI of women with PCOS ranged from 26.4 to 39.9 was
typically greater than women without PCOS which ranged 20.5 to 27.9.

@@

For mean REE the model including the studies noted with possible
reporting issues that were not resolved\textsuperscript{13,35--37} still
resulted in qualitatively similar inferences suggesting little
difference between control and PCOS conditions with mean point estimate
of 15 kcal/day with a 95\% quantile interval ranging from -69 kcal/day
to 95 kcal/day. This was similar for the standard deviation of REE with
the model including these studies resulting in a contrast ratio between
the control and PCOS conditions with a mean point estimate of 1.21 and a
95\% quantile interval ranging from 0.68 to 2.13.

@@

For mean REE the model which included only baseline REE measurements
from studies involving interventions\textsuperscript{34,53,54}, in
addition to other cross-sectional studies, also resulted in
qualitatively similar inferences suggesting little difference between
control and PCOS conditions with mean point estimate of 32 kcal/day with
a 95\% quantile interval ranging from -46 kcal/day to 118 kcal/day. This
was also the case for the standard deviation of REE with this model
resulting in a contrast ratio between the control and PCOS conditions
with a mean point estimate of 0.97 and a 95\% quantile interval ranging
from 0.71 to 1.36.

@@

These findings challenge the widely held belief that PCOS is inherently
associated with a slower metabolism\textsuperscript{@ 25}, predisposing
women with PCOS to weight gain. This belief largely stems from a single
but influential 2009 study from Georgopoulos et al.~that reported a
significantly reduced BMR in women with PCOS\textsuperscript{13}, which
has been widely cited and reinforced in both academic and clinical
contexts. However, as we have noted, this study along with
others\textsuperscript{35--37} have numerous reporting errors which led
us to drop them from our present analysis (though sensitivity analysis
including them did not alter our overall conclusions). The mistaken
belief that REE is lower for those with PCOS may have mistakenly lead to
recommendations centred on slightly more severe calorie restriction to
achieve weight-loss goals, compared with recommendations for the general
population, as primary management strategies for women with
PCOS\textsuperscript{20}. Recognising that there may be minimal
differences in REE between women with PCOS and those without can inform
both clinical and public practices, potentially leading to a shift in
focus away from a requirement for more severe caloric restriction as a
primary method of treatment towards more comprehensive, individualised,
and psychologically safe approaches to care\textsuperscript{25}.

>>>>

A recent scoping review has highlighted the inconsistency in research
examining energy expenditure in women with PCOS\textsuperscript{30}.
However, to our knowledge there has been no attempts to systematically
review, and quantitatively synthesise via meta-analysis, studies of REE
in women with, and without, PCOS. Therefore, to estimate and describe
the magnitude of difference in REE between women with and without PCOS,
we completed a systematic review and meta-analysis of studies reporting
REE in these populations.

@@

This systematic review and meta-analysis was pre-registered on PROSPERO
(\href{https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434}{CRD42024601434})
initially on the \(3^{rd}\) of December 2024 and performed in accordance
with the Preferred Reporting Items for Systematic Reviews and
Meta-Analyses (PRISMA) statement guidelines\textsuperscript{31}. The
PRISMA flow diagram reported below (Figure~\ref{fig-prisma}) was
produced using the \texttt{PRISMA2020} R package and Shiny
app\textsuperscript{32}. The primary aim of this review was to examine
the descriptive question ``Does resting energy expenditure (REE) differ
between women with and without polycystic ovary syndrome (PCOS)?''. We
summarise and describe the studies in addition to quantitatively
synthesising their results via meta-analysis.

@@

The condition being studied was PCOS and we included observational
cross-sectional design studies, in addition to intervention studies
where REE was reported for the population (and if present, the
comparator i.e., women without PCOS) condition of interest. For clarity,
studies of any design were included if they reported the REE using the
methods indicated for a sample of adult women with PCOS and who were
otherwise healthy. This included both studies with and without samples
of healthy control women without PCOS. As detailed in the statistical
analysis section below, a Bayesian model with informative priors based
on normative data for REE in healthy women without PCOS was included to
provide control information indirectly where this was missing. The use
of such priors is an efficient tool for incorporating historical
information on a particular population in a conservative
manner\textsuperscript{33}.

@@

For each arm, and observation time point if multiple observations
reported (e.g., before and after an intervention), depending on what was
reported by the authors we extracted the means, medians, standard
deviations, standard errors, lower and upper range values, and
interquartile range for the unadjusted and/or body mass adjusted and/or
fat free mass adjusted REE values. Where REE values adjusted for body
mass and/or fat free mass were reported we used the reported body mass
and/or fat free mass mean values for that arm to convert them to
unadjusted REE values (i.e., multiplied them by body mass and/or fat
free mass mean values). Where means and/or standard deviations were
missing the latter were either calculated from standard errors and
sample size, or all both were estimated from lower and upper range,
interquartile range, median, and sample size depending on the available
information using the methods of Wan et al.\textsuperscript{34}.
Further, where missing, height/body mass/BMI where estimated based on
the reported means. The units of measurement for which REE was extracted
and all REE values were converted to kcal/day. In one
case\textsuperscript{35} REE was reported relative to body mass and the
unadjusted values were no longer available (confirmed by the authors).
As such, in this case we used the mean body mass to convert back to
estimated REE unadjusted.

@@

During data extraction it was noted that several studies from the same
lab/research group\textsuperscript{13,36--38} contained a number of
discrepancies that seemed to be possible reporting errors. This
included, based on taking the authors results as written, standard
errors that implied impossible or at least incredibly unlikely standard
deviations, and discrepancies in sample size reporting throughout for
most variables without explanation or where this was explained the
sample sizes were discrepant with the text Further, data was not
reported for the healthy control women without PCOS in three of the
studies\textsuperscript{36--38}, and REE was reported as an ``adjusted''
value whereby
\(\textrm{REE}_{adjusted}=\textrm{REE}_{group~mean}+(\textrm{REE}_{adjusted}−\textrm{REE}_{predicted})\)
and the \(\textrm{REE}_{predicted}\) was obtained by substituting the
individual lean body mass, fat mass, gender, and age in the linear
regression equation generated by the data of all patients. In
correspondence with the senior author we were unable to clarify the
reporting discrepancies as the person responsible for the data/results
was no longer contactable. The original data were also no longer
available and so we could not calculated the unadjusted REE.

@@

Statistical analysis of the data extracted was be performed in R, (v
4.3.3; R Core Team, https://www.r-project.org/) and RStudio (v
2023.06.1; Posit, https://posit.co/). All code utilised for data
preparation, transformations, analyses, plotting, and reporting are
available in the corresponding GitHub repository
\url{https://github.com/jamessteeleii/pcos_ree_meta}. We cite all
software and packages used in the analysis pipeline using the
\texttt{grateful} package\textsuperscript{39} which can be seen here:
\url{https://github.com/jamessteeleii/pcos_ree_meta/blob/main/grateful-report.pdf}.
The statistical analysis plan was linked in our pre-registration
(PROSPERO:
\href{https://www.crd.york.ac.uk/PROSPERO/view/CRD42024601434}{CRD42024601434})
and available at the accompanying GitHub repository. Any deviations from
the pre-registration are noted below.

Given our research question our analysis was aimed at parameter
estimation\textsuperscript{40} within a Bayesian meta-analytic
framework\textsuperscript{41}. For all analyses model parameter
estimates and their precision, along with conclusions based upon them,
are interpreted continuously and probabilistically, considering data
quality, plausibility of effect, and previous literature, all within the
context of each model. The \texttt{renv} package\textsuperscript{42} was
used for package version reproducibility and a function based analysis
pipeline using the \texttt{targets} package\textsuperscript{43} was
employed (the analysis pipeline can be viewed by downloading the R
Project and running the function \texttt{targets::tar\_visnetwork()}).
Effect sizes and their variances were all calculated using the
\texttt{metafor} packages \texttt{escalc()}
function\textsuperscript{44}. The main package
\texttt{brms}\textsuperscript{45} was used in fitting all the Bayesian
meta-analysis models. Prior and posterior draws were taken using
\texttt{marginaleffects}\textsuperscript{46} and
\texttt{tidybayes}\textsuperscript{47} packages. All visualisations are
created using \texttt{ggplot2}\textsuperscript{48}, \texttt{tidybayes},
and the \texttt{patchwork}\textsuperscript{49} packages.

@@

We adopted an arm-based multiple condition comparison (i.e., network)
type model given that the studies included had arms of women with PCOS
both with, and without, a non-PCOS control arm\textsuperscript{50}, and
also in some cases multiple observations of REE in the different arms
included in the study (for example, where an intervention was conducted
and pre- and post-intervention REE was reported). In typical
contrast-based meta-analyses data is limited to the effect sizes for
paired contrasts between arms and thus studies that include both arms
(i.e., relative effects between non-PCOS control vs PCOS arms); however,
in arm-based analyses the data are the absolute effects within each arm
and information is borrowed across studies to enable both within
condition absolute, and between condition relative contrasts to be
estimated. We made use of historical information regarding REE in
healthy control women without PCOS by setting informative priors based
on meta-analysis of large scale studies reporting normative data for REE
in this population. This was included to provide indirect control
information where it was missing from particular studies. The use of
historical priors like this is an efficient tool to incorporate
historical information about a particular population in a conservative
manner in meta-analyses\textsuperscript{33}. From this model we focus on
reporting the global grand mean estimate for the fixed between condition
relative contrast for non-PCOS control vs PCOS arms as our primary
estimand of interest (i.e., \(\beta_1\) in both mean and standard
deviation models). We examined both raw mean REE (i.e., the absolute
mean REE in kcals per day for each arm) in addition to the between
person standard deviation in REE (i.e., the absolute standard deviation
in REE in kcals per day for each arm). Both models were multilevel in
that they included nested random intercepts for both study and arm
within study. In addition, and in deviation from our pre-registration,
we also included lab as a random intercept as in some cases we had
multiple studies from the same lab or research group. Lastly, the
inclusion of a random intercept for each effect size was accidentally
omitted from our pre-registration, and so this is also included in the
model.

@@

\[
\small
\begin{aligned}
\beta_0 &\sim \text{Student-}t(3,\, 1441.81,\, 84.56) \\[4pt]
\beta_1 &\sim \text{Student-}t(3,\, 0,\, 200) \\[12pt]
\sigma^2_{0,\mathrm{study}} &\sim \text{Half-student-}t(3,\, 149.89,\, 82.91) \\[4pt]
\sigma^2_{0,\mathrm{lab}},\, \sigma^2_{1,\mathrm{lab}},\,
\sigma^2_{1,\mathrm{study}},\, \sigma^2_{0,\mathrm{arm}}, \sigma^2_{0,\mathrm{effect}} &\sim \text{Half-student-}t(3,\, 0,\, 112.4) \\[12pt]
\mathrm{corr}_{\mathrm{lab}} &\sim \mathrm{LKJ}(1) \\[4pt]
\mathrm{corr}_{\mathrm{study}} &\sim \mathrm{LKJ}(1)
\end{aligned}
\] where the prior for \(\beta_{0}\), which corresponded to the model
intercept and mean REE in the control condition, was set based on
meta-analysis of the mean REEs for women from two large studies of
healthy people\textsuperscript{51,52} though set with a conservative
degrees of freedom for the \(\textrm{Student-}t\) distribution.The
random intercept \(\sigma^2_{0,study}\) was set similarly to this. The
prior for the fixed effect \(\beta_{1}\), reflecting the difference
between control and PCOS conditions was set based on a wide range of
possible values considering the minimum and and maximum values of the
ranges reported in the two studies noted (i.e., 2492 - 908 = 1584). We
then set a prior that permits values approximately across this range of
values with the majority of it's mass centred around zero. The remaining
random effects were set based on the default weakly regularising priors
for \texttt{brms} and scaled to the expected response values using a
\(\textrm{half-student-}t\) distribution with 3 degrees of freedom and
\(\mu=0\), and both correlation matrices \(\textrm{corr}_{lab}\) and
\(\textrm{corr}_{study}\) were set with an \(\textrm{LKJcorr(1)}\)
distribution.

@@

where the prior for \(\beta_{0}\), which corresponded to the model
intercept and standard deviation of REE in the control condition, was
again set based on meta-analysis of the standard deviation of REEs for
women from two large studies of healthy people\textsuperscript{51,52}
though set with a conservative degrees of freedom for the
\(\textrm{Student-}t\) distribution.The random intercept
\(\sigma^2_{0,study}\) was set similarly to this. The prior for the
fixed effect \(\beta_{1}\), reflecting the difference between control
and PCOS conditions was set based on a wide range of possible values.
Given that in many cases of variables in the field there is an
approximate relationship of \textasciitilde1 for the natural logarithm
of the standard deviation conditioned upon the natural logarithm of the
mean\textsuperscript{53} we set this prior to reflect the range of
differences on the on the log scale (i.e., \(\log(1584)\)). We then set
a prior that permits values approximately across this range of values
with the majority of it's mass centred around zero. The prior for the
fixed effect \(\beta_{2}\), reflecting the relationship between the
natural logarithm of the mean REE with the natural logarithm of the
standard deviation of REE, was set it to be weakly informative centred
on zero with a wide scale to indicate uncertainty in this outcome
specifically despite the typical relationship close to \textasciitilde1.
Priors for the measurement error of the natural logarithm of the mean
REE estimate \(\tilde{m}\) were again based upon meta-analysis of the
aforementioned studies, though measurement error has to be positive, as
does the corresponding standard deviation of this error, so we set these
to conservative wide \(\textrm{half-student-}t\) distributions. The
remaining random effects were set based on the default weakly
regularising priors for \texttt{brms} and scaled to the expected
response values using a \(\textrm{half-student-}t\) distribution with 3
degrees of freedom and \(\mu=0\), and both correlation matrices
\(\textrm{corr}_{lab}\) and \(\textrm{corr}_{study}\) were set with an
\(\textrm{LKJcorr(1)}\) distribution.

@@

As noted in the section above, \emph{``Studies with possible reporting
errors''}, we opted to conduct analysis with and without the inclusion
of four studies with possible reporting errors we could not
resolve\textsuperscript{13,36--38}. Thus the main models described above
were run with and without these studies, the main results are presented
without them and the sensitivity results with them are reported
separately below.

As a further sensitivity analysis, given the inclusion of some studies
with interventions in women with PCOS reporting REE at multiple
timepoints such as mid- and post-intervention\textsuperscript{35,54,55}
but lacking of control women without PCOS, we opted to also conduct
sensitivity analysis excluding these and only examining their baseline
results in addition to the cross-sectional studies. As such, the main
models described above were run without the follow-up (i.e., mid- or
post-intervention timepoints) from these studies and only using the
baseline results in addition to other cross-sectional studies.

@@

Note, the numbers in this section exclude the four studies previously
noted with possible reporting issues that were unresolved. These studies
are however summarised in fully in the descriptive characteristics table
in the online supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table} where some of the possible reporting errors are seen in the
standard deviations reported or calculated e.g., the standard deviation
calculated for body mass of women with PCOS in Saltamavros et
al.\textsuperscript{37}. The PRISMA flow diagram
(Figure~\ref{fig-prisma}) also includes all studies identified including
the four with possible errors.

@@

Studies were carried out in multiple countries: Brazil (k = 3) ,and USA
(k = 3 studies), Australia, Cameroon, Canada, Italy, Sweden, Turkey, UK
(all k = 1 study). The four studies with noted reporting errors were
carried out in Greece. A total of 11 studies used the Rotterdam criteria
(or a modified version there of) for diagnosing
PCOS\textsuperscript{16,17,35,54--61}. One study used the 1990 National
Institutes of Health criteria\textsuperscript{62}, and two
studies\textsuperscript{18,56} diagnosed PCOS via the presence of
oligomenorrhea or amenorrhea alongside additional criteria including
plasma androgen levels, hirsutism or polycystic ovaries on ultrasound
scanning. All but one study\textsuperscript{62}, which used doubly
labelled water, measured REE using indirect calorimetry. The specific
devices reported by these studies are included in the online
supplementary materials (see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}). We originally considered in our pre-registration that, given
sufficient data, we would compare sub groups of women with PCOS who did
and did not have accompanying insulin resistance. However, based on the
metabolic health variables reported in studies with this information
(see
\href{https://html-preview.github.io/?url=https://github.com/jamessteeleii/pcos_ree_meta/raw/refs/heads/main/tables/descriptives_table.html}{Descriptives
Table}) and, where available, considering primary criteria of either
homeostatic model assessment of insulin resistance (HOMA-IR) \(\geq2.5\)
or secondary criteria including fasting insulin \(>12\mu \textrm{U/mL}\)
or fasting glucose \(\geq100\textrm{mg/dL}\), all groups of women with
PCOS in the included studies would be considered to have insulin
resistance. Mean age of women with PCOS in these studies ranged from 23
to 33 which was similar to the control women without PCOS ranging from
23 to 30. Across those studies where BMI was reported or it was possible
to estimate the mean BMI of women with PCOS ranged from 26.4 to 39.9 was
typically greater than women without PCOS which ranged 20.5 to 27.9.

@@

For mean REE the model including the studies noted with possible
reporting issues that were not resolved\textsuperscript{13,36--38} still
resulted in qualitatively similar inferences suggesting little
difference between control and PCOS conditions with mean point estimate
of 15 kcal/day with a 95\% quantile interval ranging from -69 kcal/day
to 95 kcal/day. This was similar for the standard deviation of REE with
the model including these studies resulting in a contrast ratio between
the control and PCOS conditions with a mean point estimate of 1.21 and a
95\% quantile interval ranging from 0.68 to 2.13.

@@

For mean REE the model which included only baseline REE measurements
from studies involving interventions\textsuperscript{35,54,55}, in
addition to other cross-sectional studies, also resulted in
qualitatively similar inferences suggesting little difference between
control and PCOS conditions with mean point estimate of 32 kcal/day with
a 95\% quantile interval ranging from -46 kcal/day to 118 kcal/day. This
was also the case for the standard deviation of REE with this model
resulting in a contrast ratio between the control and PCOS conditions
with a mean point estimate of 0.97 and a 95\% quantile interval ranging
from 0.71 to 1.36.

@@

These findings challenge the widely held belief that PCOS is inherently
associated with a slower metabolism\textsuperscript{@ 25}, predisposing
women with PCOS to weight gain. This belief largely stems from a single
but influential 2009 study from Georgopoulos et al.~that reported a
significantly reduced BMR in women with PCOS\textsuperscript{13}, which
has been widely cited and reinforced in both academic and clinical
contexts. However, as we have noted, this study along with
others\textsuperscript{36--38} have numerous reporting errors which led
us to drop them from our present analysis (though sensitivity analysis
including them did not alter our overall conclusions). The mistaken
belief that REE is lower for those with PCOS may have mistakenly lead to
recommendations centred on slightly more severe calorie restriction to
achieve weight-loss goals, compared with recommendations for the general
population, as primary management strategies for women with
PCOS\textsuperscript{20}. Recognising that there may be minimal
differences in REE between women with PCOS and those without can inform
both clinical and public practices, potentially leading to a shift in
focus away from a requirement for more severe caloric restriction as a
primary method of treatment towards more comprehensive, individualised,
and psychologically safe approaches to care\textsuperscript{25}.

The current study suggests that REE may not a barrier to weight
regulation in PCOS given small group-level differences between women
with PCOS and those without (1472 kcal/day {[}95\%QI:1359 kcal/day to
1587 kcal/day{]} versus 1442 kcal/day {[}95\%QI:1334 kcal/day to 1553
kcal/day{]}, respectively). If anything, REE may be slightly higher in
women with PCOS compared to healthy women without PCOS. BMI of women
with PCOS ranged from 26.4 to 39.9 was typically greater than women
without PCOS which ranged 20.5 to 27.9 and this may explain the slightly
greater REE in the former group. However, for those studies where we
could extract or estimate BMI, our additional exploratory models
adjusted for this similarly showed little difference in REE (20 kcal/day
{[}95\%QI: -96 kcal/day to 140kcal/day{]}) between women with PCOS and
those without. However, one of the studies included in our
analysis\textsuperscript{63} reported that, whilst there was little
difference in unadjusted REE, when REE was adjusted for fat-free mass it
was lower in women with PCOS suggesting the potential importance of
fat-free tissue in energy regulation. Yet, for those studies where
fat-free mass was reported, our additional exploratory models adjusted
for this similarly showed little difference in REE (-9 kcal/day
{[}95\%QI: -195 kcal/day to 168kcal/day{]}) between women with PCOS and
those without. As such, even adjusted for both BMI and fat-free mass,
there seems to be little difference in REE between women with, and
without, PCOS. Yet, a recent systematic review of mechanisms for
metabolic dysfunction has reported excess androgen drives metabolic
issues within adipose tissue and muscle tissue contributing to
complications like obesity and insulin resistance\textsuperscript{64}.
Taken together, these findings highlight that factors beyond REE and
typical correlates of this including BMI or fat-free mass, such as
hormonal and tissue-specific metabolic effects, may play a more
significant role in weight regulation challenges in women with PCOS.
